State of the Practice for Medical Imaging Software

### 2022-10-18 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Edits, mostly to the recommendations section

The above discussion shows that, taken together, MI projects fall somewhat short
of recommended best practices for research software.  However, MI software is
not alone in this.  Many, if not most, research projects fall short of best
practices.  A gap exists in scientific computing development practices and
software engineering recommendations \citep{Storer2017, Kelly2007,
OwojaiyeEtAl2021_CSE}. \citet{JohansonAndHasselbring2018} observe that the
state-of-the-practice for SCS in industry and academia does not incorporate
state-of-the-art SE tools and methods.  This causes sustainability and
reliability problems \citep{FaulkEtAl2009}. Rather than benefit from capturing
and reusing previous knowledge, projects waste time and energy ``reinventing the
wheel'' \citep{deSouzaEtAl2019}.

@@

We answer research question \rqref{RQ_CompareMethodologies} by comparing the
principles, processes and methodologies used for MI software to what can be
gleaned from the literature on research software in general. In our interviews
with developers the responses about development model were vague, with only two
interviewees following a definite development model. In some cases the
interviewees felt their process was similar to an existing development model.
Three teams (about 38\%) either followed agile, or something similar to agile.
Two teams (25\%) either followed a waterfall process, or something similar.
Three teams (about 38\%) explicitly stated that their process was undefined or
self-directed.

Our observations of an informally defined process, with elements of agile
methods, matches what has been observed for research software in general.
Scientific developers naturally use an agile philosophy \citep{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \citep{Kelly2013}, or a knowledge acquisition driven
process \citep{Kelly2015}.  A waterfall like process can work for research
software \citep{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if a rationale design
process were followed \citep{parnas1986rational}.

@@

\citet{PintoEtAl2018} lists some pain points that did not come up in our
conversations with MI developers: interruptions while coding, scope bloat, lack
of user feedback, hard to collaborate on software projects, and aloneness.
\citet{WieseEtAl2019} also mention two research software pain points that did
not explicitly arise in our interviews: reproducibility, and software scope
determination.  To the list of pain points not discussed for MI, our study of
LBM software \citep{SmithEtAl2022} adds lack of software experience for the
developers, technical debt, and documentation. We did not observe any pain
points for MI that were not also observed for LBM. From the pain points
mentioned above, although the topics of reproducibility and technical debt did
not come up in our MI interviews, these two topics were discussed as part of the
discussion of software qualities, as covered at the end of this section.
Although previous studies show pain points that were not mentioned by MI
developers, we cannot conclude that these pain points are not relevant for MI
software development, since we only interviewed nine developers for about an
hour each.

@@

\begin{itemize}
\item Using open datasets, like the datasets for the liver \citep{BilicEtAl2019}
and brain \citep{MenzeEtAl2015} tumor segmentation benchmarks.
\item If (part of) the team belongs to a medical school or a hospital, using the
datasets they can access;
\item If the team has access to MRI scanners, self-building sample images for
testing;
\item If the team has connections with MI equipment manufacturers, asking for
their help on data format problems;
\end{itemize}

@@

\item[Q\refstepcounter{qnum}\theqnum \label{Q_Reproducibility}:]
\textbf{Reproducibility:}  Although the MI developers did not mention
reproducibility explicitly as a pain point, they did mention the need to improve
documentation.  Good documentation doesn't just address the pain points of lack
of developer time (\ppref{P_LackDevTime}), technology hurdles
(\ppref{P_TechnologyHurdles}), usability \ppref{P_Usability}, and
Maintainability.  Documentation is also necessary for reproducibility. The
challenges of inadequate documentation are a known problem for research software
\citep{PintoEtAl2018, WieseEtAl2019} and for non-research software
\citep{LethbridgeEtAl2003}.

@@

\subsection{Move To Web Applications} \label{sec_recommendations_tech_stack}

Section~\ref{painpoints} describes technology hurdles
(\ppref{P_TechnologyHurdles}) that motivate considering the use of web
applications. Here we give further advice to help with deciding whether to adopt
a web application. The decision will be based on whether, on balance, the web
application improves the four factors identified by developers: compatibility,
maintainability, performance, and security. To enable decision-making, a team
will need to prioritize between these factors, based on the team's objectives
and experience. The suggestions are intended to provide ideas and avenues for
exploration; a web application might not be the right fit for all projects and
all teams.

@@

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.

\item \textbf{Web applications can potentially deliver high performance.} Web
applications with backend servers may perform even better than native
applications. If a team needs to support lower-end computers, it is good to use
back-end servers for heavy computing tasks.

\item \textbf{Backend servers can have low costs.} Serverless solutions from
major cloud service providers may be worth exploring. Serverless solutions still
use a server, but the server provider only charges the team when they use the
server. The solution is event-driven, and costs the team by the number of
requests processed. Thus, serverless can be very cost-effective for less
intensively used functions.

@@

\subsection{Recommendations on Enriching the Testing Datasets} 
\label{sec_recommendations_testing_dataset}

As described in Section~\ref{painpoints} technology hurdles
(\ppref{P_TechnologyHurdles}) make it difficult for software development teams
to access real-world medical imaging datasets. This problem restricts their
capability and flexibility for testing. We provide some suggestions as follows:

@@

\item \textbf{Search for open data sources.} In general, there are many open MI
datasets.  For instance, there are
\href{https://nihcc.app.box.com/v/ChestXray-NIHCC}{Chest X-ray Datasets} by
National Institute of Health \citep{WangEtAl2017},
\href{https://www.cancerimagingarchive.net/}{Cancer Imaging Archive}
\citep{PriorEtAl2017}, and \href{https://medpix.nlm.nih.gov/home}{MedPix} by
National Library of Medicine \citep{Smirniotopoulos2014}. A team developing MI
software should be able to find more open datasets according to their needs.

@@

\item \textbf{Establish community collaboration in the domain.} During our
interviews with developers in the MI domain, we heard many stories of asking for
supports from other professionals or equipment manufacturers. However, we
believe that broader collaboration between development teams can address this
problem better. Some datasets are too sensitive to share, but if the community
has some kind of ``group discussion'', teams can better express their needs, and
professionals can better offer voluntary support for testing. Ultimately, the
community can establish a nonprofit organization as a third-party, which
maintains large datasets, tests OSS in the domain, and protects privacy.

@@

Except for \citet{ThielEtAl2020}, the research software guidelines that we
consulted do not mention linters.  However, we believe that linters have the
potential to improve code quality at a relatively low cost.  A linter is a tool
that statically analyzes code to find programming errors, stylistic
inconsistencies and suspicious constructs \citep{Wikipedia2022_Lint}. Linters
can be used to spot check code files, or even better as part of a continuous
integration system, as discussed in Section~\ref{Sec_ContinuousIntegration}.

A sample of the potential benefits of linters include: finding memory leaks,
finding potential bugs, standardizing code with respect to formatting, improving
performance, removing silly errors before code reviews, and catching potential
security issues \citep{SourceLevel2022_Lint}.  Linters are available for most
popular programming languages.  For instance Python has the options of PyLint,
flake8 and Black \citep{Zadka2018}.

Linters can address the pain points for LBM developers.  For instance, a linter
can decrease the amount of development time (\ppref{P_LackDevTime}) by
decreasing the number of mundane mistakes programmers have to catch.  Since the
linter can include rules that capture the wisdom of senior programmers, it can
help newer developers avoid common mistakes. Although there are not many linters
for checking parallel programming mistakes (\ppref{P_TechnologyHurdles}),
Parallel Lint is an option for OpenMP. Consistently using a linter can guard
against some technical debt (\ppref{Q_Maintainability}). Although linters are
tools for code analysis, similar ideas can be applied to ``lint'' documentation
to ensure adherence to basic rules.  The use of tools to check documentation is
a partial explanation for the relatively higher quality of statistical tools
that are part of the Comprehensive R Archive Network (CRAN)
\citep{SmithEtAl2018_StatSoft}.

\subsection{Conduct Rigorous Peer Reviews} \label{Sec_PeerReview}

Peer review is a strategy already adopted by some LBM developers to improve
confidence in their software, but the benefits of peer review can be improved by
making its application more rigorous. \citet{Jones2008} shows that rigorous
inspection finds 60-65\% of latent defects on average, and often tops 85\% in
defect removal efficiency.  By way of comparison, most forms of testing average
between 30 and 35\% for defect removal efficiency \citep{EbertAndJones2009,
Jones2008}.  A formal code inspection involves asking the reviewer to either
follow a review checklist (check consistency of variable names, look for
terminating loops, etc.), or perform specific review tasks (like summarize the
purpose of the code, create a data dictionary for a given module,
cross-reference the code to the technical manual, etc.) The task based
inspection approach has been effectively used for research software, as
described by \citet{KellyAndShepard2000}. Task based inspection is an ideal fit
with an issue tracking system, like GitHub.  The review tasks can be issues, so
that they can be easily assigned, monitored and recorded.  Other potential
issues for the tracker include assigning junior team members to test
installation instructions and getting-started tutorials.

Rigorous peer review addresses the same pain points as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \ppref{Q_Maintainability}. Peer review addresses these points by efficiently
searching for defects and problems.  Finding misunderstandings in of how the
code implements the required theory improves the software's correctness
(\ppref{P_Correctness}).

Peer review --  One of the developers (ESPResSo) that was interviewed noted that
an ad hoc peer review process is used to assess major changes and additions.
Using peer review (also called technical review) matches with recommended
practice for research software \citep{HerouxEtAl2008, Givler2020, OrvizEtAl2017,
USGS2019}.

\subsection{Follow Advice From the Open-Source Community to Grow the Number of Contributors} \label{Sec_GrowContributors}

In our interviews with developers we heard they would like to have additional
software contributors.  More developers would help with addressing the lack of
development time pain point (\ppref{P_LackDevTime}).  In investigating advice on
increasing the number of contributors, we found that the advice usually starts
with following best practices, including providing artifacts like those
discussed in Section~\ref{Sec_CompareArtifacts}, such as a contributor
guidelines, code style guidelines, a clear code of conduct and high quality code
and documentation. This advice is logical, but in our assessment of LBM software
we found examples of projects that already follow many best practices, but still
have small development teams. Attracting developers apparently requires more
than just building a good product.

Some potential additional ideas for growing the number of contributors are
as follows:

\begin{itemize}
	\item Clearly identify issues that are an appropriate starting points for
	new developers \citep{Garcia2016, Jalan2016, Proffitt2017}.
	\item Given that in most projects developers were once users
	\citep{McQuaid2018}, recruit future developers from the current set of
	users.
	\item Create templates for pull requests and issues \citep{Jalan2016}.
    \item Welcome all kinds of contributions, not just code.  Non-code
    contributions include documentation, fixing typos, issue reporting and test
    cases \citep{Jalan2016, Proffitt2017}.
    \item Reward and recognize new contributors, via small rewards like stickers
    or shirts, or even just a simple shoutout or mention in a blog post or on
    social media \citep{Jalan2016, Proffitt2017}.
    \item Recognize that open source is more about people than it will ever be
    about code \citep{Jalan2016}.
    \item Look beyond just recruiting online and seek new developers at
    conferences or other user meet ups \citep{Garcia2016}.
    \item Follow the advice of \citet{Kuchner2012} to adapt ideas from marketing
    to promote science.  Specific ideas that could be applied to open source
    projects include the following: 
    \begin{itemize}
		\item Storytelling to motivate interest in the project, where the story
		is a sequence of events and pauses for reflection \citep[p.\
		21--22]{Kuchner2012}.
		\item Invest effort in building relationships with users and with potential future developers.
		\item Brand your project with what makes it unique and special.
	\end{itemize}
	\item Consider combining existing projects to pool collective resources, and
	reduce competition between different open source projects.
\end{itemize}

@@

To address technical debt (\ppref{Q_Maintainability}), the top LBM developers show how
modularization can be used to design research software for future change.
Although the advice to modularize research software to handle complexity is
common \citep{WilsonEtAl2014, StewartEtAl2017, Storer2017}, specific guidelines
on how to divide the software into modules is less prevalent.  Not every
decomposition is a good design for supporting change, as shown by
\citet{Parnas1972a}.  A design with low cohesion and high coupling \citep[p.\
48]{GhezziEtAl2003} will make change difficult. Especially in research software,
where change is inevitable, designers need to produce a modularization that
supports change. Ocean modelling software is currently feeling the pain of
not emphasizing modularization in legacy code \citep{JungEtAl2022}.

Interviewed LBM developers highlighted cases where their modularizations
anticipated future changes.  The developer of pyLBM mentioned that the
geometries and models of their system had been ``decoupled'', using abstraction
and modularization of the source code, to make it ``very easy to add [new]
features''.  The pyLBM design allows for independent changes to the geometry and
the model.  We also learned that the package pyLBM redeveloped data structures
to ease future changes. The developer of TCLB noted that their design allows for
the addition of some LBM features, but changes to major aspects of the system
would be difficult. For example, ``implementing a new model will be an easy
contribution'', but changes to the ``Cartesian mesh … will be a nightmare''. The
design of TCLB highlights that not every conceivable change needs to be
supported, only the likely changes.

As the LBM developers illustrate, they accomplish design for change by first
identifying likely changes, either implicitly or explicitly, and second by
hiding each likely change behind a well-defined module interface.  Although it
is unclear whether the developers are aware of Parnas's work, the approach
mirrors his recommendations \citep{Parnas1972a}. \citet{SmithEtAl2022} (Section
Documentation as Part of the Development Process) lists ideas for how to
document the design, including the likely changes, so that they are more visible
to others.

@@

From \cite{RinehartEtAl2015}, an assurance case is ``[a] reasoned and compelling
argument, supported by a body of evidence, that a system, service or
organization will operate as intended for a defined application in a defined
environment.''  Assurance cases have been successfully employed for safety
critical systems, but using this technique for SC software is a new idea.  The
Goal Structuring Notation (GSN) \cite{Spriggs2012} seems like a good framework
for initial investigation.

Scientific software, such as medical software, is often subject to
standardization and regulatory approval. While applying such approvals and
standards has had a beneficial effect on system quality, it does not provide
good tracking of the development stages, as the compliance with the standards
is mostly checked after the system development. Once a system is implemented,
its documentations must be approved by the regulators. This process is lengthy
and expensive. In contrast, assurance case development usually occurs in
parallel with the system construction, resulting in a traceable, detailed
argument for the desired property.  Moreover, since the assurance case is
integrated into the system construction, the domain experts are involved from
the start.

Putting the argument in the hands of the experts means that they will work to
convince themselves, along with the regulators.  They will use the expertise
that the regulators do not have and they will see the value of documentation (as
mentioned in section on pain points with methods to mitigate); they will be
engaged.  This engagement will hopefully help bridge the current chasm between
software engineering and scientific computing~\cite{Kelly2007}, by motivating
scientists toward documentation and correcting the problem of software engineers
failing to meet scientists' expectations~\cite{Segal2008}.  Significant
documentation will still likely be necessary, but now the developers control the
documentation.  What is created will be relevant and necessary.

For SCS a top level claim could read ``Program X delivers correct outputs when
used for its intended use/purpose in its intended environment.''  The next step
would be to decompose this claim into sub-claims that will be easier to prove.
The sub-claims will likely also be further divided until the bottom of the
graph, where the measurable evidence is provided.  Typical evidence will consist
of documents, expert reviews, test case results etc.

Using the specific example of preparing an assurance case for pre-existing
medical image analysis software (3dfim+)~\cite{SmithEtAl2018_ICSEPoster}, the
top level can be decomposed into four sub-goals, as shown in
Figure~\ref{TopGoal}.  This example follows the same pattern as used for medical
devices~\cite{Wassyng2015}.  The first sub-goal (GR) argues for the quality of
the documentation of the requirements (SRS).  The second sub-goal (GD) says that
the design complies with the requirements and the third proposes that the
implementation also complies with the requirements.  The fourth sub-goal (GA)
claims that the inputs to 3dfim+ will satisfy the operational assumptions, since
we need valid input to make an argument for the correctness of the output.

\begin{figure*}[!h]
\centering
\includegraphics[width=1.0\textwidth]{./figures/TopGoal.pdf}
\caption{Top Goal of the assurance case and its sub-goals}
\label{TopGoal}
\end{figure*}

Preparing an assurance case for the pre-existing 3dfim+ software justifies the
value of assurance cases for the certification of SCS~\cite{SmithEtAl2018}.
Although no errors were found in the output of the existing software, the rigour
of the proposed approach did lead to finding ambiguities and omissions in the
existing documentation, such as missing information on the coordinate system
convention.  In addition, a potential concern for the software itself was
identified from the GA argument: running the software does not produce any
warning about the obligation of the user to provide data that matches the
parametric statistical model employed for the correlation calculations.

Potential objectives for research on assurance cases applied to SCS are as
follows:

\begin{enumerate}
\item Creation of an assurance case template will require looking at a variety
  of SCS projects, which will require engaged project partners; therefore, the
  initial approach should be to target domains where certification is required,
  or domains where verification is critical, but challenging, like the High
  Energy Physics (HEP). \wss{reference the paper about domain specific
  methodology - maybe I already did?}
\item Evidence is provided at the bottom of an assurance case, but what is
  the best form for this evidence?  As mentioned for the SRS, we need an
  approach for verification.  We also need an approach to verify the other
  software artifacts, the traceability between them, the team of developers, the
  team of reviewers, etc.  The question for the future is what combination of
  tools and processes will provide convincing evidence?
\item Creation of assurance cases is challenging with existing tools.  In the
  future, the goal should be automatic generation (see
  \citet{Smith2018}, Section SecAutoGenerate) of the visual assurance case documentation from
  the arguments and evidence
\item Creating a revelation on the value of documentation \citep{Smith2018} may
   require a pseudo adversarial approach.  SCS developers should be explicitly
   challenged to present their verification efforts so that their work can be
   independently verified.  Although SCS developers have developed many
   successful theories, techniques, and testing procedures for verification, the
   evidence is generally presented in an ad hoc manner.  Developers should be
   asked to connect all the pieces of their evidence in a coherent argument.  As
   a bonus, the act of creating the assurance case may also lead developers to
   discover subtle edge cases, which would not have been noticed with a less
   rigorous and systematic approach.

@@

@@

>>>>

The above discussion shows that, taken together, MI projects fall somewhat short
of recommended best practices for research software.  However, MI software is
not alone in this.  Many, if not most, research projects fall short of best
practices.  A gap exists in scientific computing development practices and
software engineering recommendations \citep{Storer2017, Kelly2007,
OwojaiyeEtAl2021_CSE}. \citet{JohansonAndHasselbring2018} observe that the
state-of-the-practice for research software in industry and academia does not incorporate
state-of-the-art SE tools and methods.  This causes sustainability and
reliability problems \citep{FaulkEtAl2009}. Rather than benefit from capturing
and reusing previous knowledge, projects waste time and energy ``reinventing the
wheel'' \citep{deSouzaEtAl2019}.

@@

We answer research question \rqref{RQ_CompareMethodologies} by comparing the
principles, processes, and methodologies used for MI software to what can be
gleaned from the literature on research software in general. In our interviews
with developers the responses about development model were vague, with only two
interviewees following a definite development model. In some cases the
interviewees felt their process was similar to an existing development model.
Three teams (about 38\%) either followed agile, or something similar to agile.
Two teams (25\%) either followed a waterfall process, or something similar.
Three teams (about 38\%) explicitly stated that their process was undefined or
self-directed.

Our observations of an informally defined process, with elements of agile
methods, matches what has been observed for research software in general.
Scientific developers naturally use an agile philosophy \citep{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \citep{Kelly2013}, or a knowledge acquisition driven
process \citep{Kelly2015}.  A waterfall like process can work for research
software \citep{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rationale design process \citep{parnas1986rational}.

@@

\citet{PintoEtAl2018} lists some pain points that did not come up in our
conversations with MI developers: interruptions while coding, scope bloat, lack
of user feedback, hard to collaborate on software projects, and aloneness.
\citet{WieseEtAl2019} also mention two research software pain points that did
not explicitly arise in our interviews: reproducibility, and software scope
determination.  To the list of pain points not discussed for MI, our study of
LBM software \citep{SmithEtAl2022} adds lack of software experience for the
developers, technical debt, and documentation. We did not observe any pain
points for MI that were not also observed for LBM. From the pain points
mentioned above, although the topics of reproducibility and technical debt did
not come up in our MI interviews, we covered these two topics as part of the
discussion of software qualities, as summarized at the end of this section.
Although previous studies show pain points that were not mentioned by MI
developers, we cannot conclude that these pain points are not relevant for MI
software development, since we only interviewed nine developers for about an
hour each.

@@

\begin{itemize}
\item Using open datasets.
\item If (part of) the team belongs to a medical school or a hospital, using the
datasets they can access;
\item If the team has access to MRI scanners, self-building sample images for
testing;
\item If the team has connections with MI equipment manufacturers, asking for
their help on data format problems;
\end{itemize}

@@

\item[Q\refstepcounter{qnum}\theqnum \label{Q_Reproducibility}:]
\textbf{Reproducibility:}  Although the MI developers did not mention
reproducibility explicitly as a pain point, they did mention the need to improve
documentation.  Good documentation does not just address the pain points of lack
of developer time (\ppref{P_LackDevTime}), technology hurdles
(\ppref{P_TechnologyHurdles}), usability \ppref{P_Usability}, and
Maintainability.  Documentation is also necessary for reproducibility. The
challenges of inadequate documentation are a known problem for research software
\citep{PintoEtAl2018, WieseEtAl2019} and for non-research software
\citep{LethbridgeEtAl2003}.

@@

\subsection{Move To Web Applications} \label{sec_webapps}

Section~\ref{painpoints} describes the pain point of technology hurdles
(\ppref{P_TechnologyHurdles}), which motivates considering the use of web
applications. Here we give further advice to help with deciding whether to adopt
a web application. The decision will be based on whether, on balance, the web
application improves the four factors identified by developers: compatibility,
maintainability, performance, and security. To enable decision-making, a team
will need to prioritize between these factors, based on their objectives and
experience. The suggestions are intended to provide ideas and avenues for
exploration; a web application will not be the right fit for all projects and
all teams.

@@

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React} and \href{https://elm-lang.org/}{Elm}.

\item \textbf{Backend servers can potentially deliver high performance.} Web
applications with backend servers may perform even better than native
applications. If a team needs to support lower-end computers, it is good to use
back-end servers for heavy computing tasks.  For backend servers where traffic
and latency is not an issue, options include
\href{https://www.django-rest-framework.org/} {Django}, \href{https://laravel.com/} {Laravel} and
\href{https://nodejs.org/en/} {Node.js}.  The advantage of Django is that it provides access to Python
libraries.  For backend servers where traffic and latency is an issue,
\href{https://github.com/gin-gonic/gin} {Gin} is an option.

\item \textbf{Backend servers can have low costs.} Serverless solutions from
major cloud service providers (like Amazon Web Services (AWS) and Google Cloud
Platform) may be worth exploring. Serverless solutions still use a server, but
the server provider only charges the team when they use the server. The solution
is event-driven, and costs the team by the number of requests processed. Thus,
serverless can be very cost-effective for less intensively used functions.

@@

\subsection{Enrich the Testing Datasets} 
\label{sec_recommendations_testing_dataset}

As described in Section~\ref{painpoints}, ensuring correctness
(\ppref{P_Correctness}) via testing can be problematic because of limited access
to real-world medical imaging datasets.  We build on the suggestions we heard
from our interviewees as follows:

@@

\item \textbf{Search for open data sources.} In general, there are many open MI
datasets.  For instance, there are
\href{https://nihcc.app.box.com/v/ChestXray-NIHCC}{Chest X-ray Datasets} by
National Institute of Health \citep{WangEtAl2017},
\href{https://www.cancerimagingarchive.net/}{Cancer Imaging Archive}
\citep{PriorEtAl2017}, \href{https://medpix.nlm.nih.gov/home}{MedPix} by
National Library of Medicine \citep{Smirniotopoulos2014}, and datasets for liver
\citep{BilicEtAl2019} and brain \citep{MenzeEtAl2015} tumor segmentation
benchmarks.  A team developing MI software should be able to find more open
datasets according to their needs.

@@

\item \textbf{Establish community collaboration in the domain.} During our
interviews with developers in the MI domain, we heard many stories of asking for
supports from other professionals or equipment manufacturers. However, we
believe that broader collaboration between development teams can address this
problem better. Some datasets are too sensitive to share, but if the community
has some kind of ``group discussion'', teams can better express their needs, and
professionals can better offer voluntary support for testing. Ultimately, the
community can establish a nonprofit organization as a third party, which
maintains large datasets, tests Open Source Software (OSS) in the domain, and
protects privacy.

@@

A linter is a tool that statically analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \citep{Wikipedia2022_Lint}.
Linters can be used as an ad hoc check for code files, but they really come into
their own when used as part of a continuous integration system, as discussed in
Section~\ref{Sec_ContinuousIntegration}. Almost none of the research software
guidelines that we consulted, summarized in Section~\ref{Sec_CompareArtifacts},
mention linters.  The one exception is \citet{ThielEtAl2020}.  Despite the lack
of mention in the guidelines, we believe that linters have the potential to
improve code quality at a relatively low cost.

Linters have the following benefits: finding potential bugs, finding memory
leaks, improving performance, standardizing code with respect to formatting,
removing silly errors before code reviews, and catching potential security
issues \citep{SourceLevel2022_Lint}. Most popular programming languages have an
accompanying linter.  For example, Python has the options of PyLint, flake8 and
Black \citep{Zadka2018}.

We recommend the use of linters because they are relatively easy to incorporate
into a developer's workflow, and they address several MI pain points
(Section~\ref{painpoints}).  For instance, linters address the lack of
development time (\ppref{P_LackDevTime}) by increasing the developer's
productive time via guarding against making frustrating, time-consuming, mundane
mistakes.  Moreover, since a linter can include rules that capture the wisdom of
senior programmers, it can help newer developers avoid common mistakes. With
respect to the technology hurdle pain point (\ppref{P_TechnologyHurdles}),
linters can assist with the move toward web applications
(Section~\ref{sec_webapps}).  For instance, ESLint in React is a pluggable
linter that lets the developer know if they have imported something and not used
it, if a function could be short-handed, if there are indentation
inconsistencies, etc. \citep{Whitehouse2018}. By insisting on code
standarization linters can reduce technical debt and thus improve
maintainability (\qref{Q_Maintainability}). Although linters are tools for code
analysis, the idea of statically checking for adherence to basic rules can be
extended to check documentation. \citet{SmithEtAl2018_StatSoft} shows how the
use of tools to enforce documentation standards partially explains the
relatively higher quality of statistical tools that are part of the
Comprehensive R Archive Network (CRAN).

\subsection{Conduct a Mix of Rigorous and Informal Peer Reviews} \label{Sec_PeerReview}

We advocate incorporating peer review into the development process, as
frequently recommended for research software \citep{HerouxEtAl2008, Givler2020,
OrvizEtAl2017, USGS2019}. In most cases a modern, lightweight review, should be
adequate.  Modern code review is informal, tool-based, asynchronous, and focused
on reviewing code changes \citep{SadowskiEtAl2018}. Managing a project via
GitHub pull requests is an example of a modern approach to reviewing code.
Software development organizations have moved to this lightweight style of code
review because of the inefficiencies of rigorous inspections
\citep{RigbyAndBird2013}.  However, for important parts of the code, developers
may benefit from mixing in a more rigorous approach.

\citet{Fagan1976} began work on rigorous review via code inspection.  Elements
of a typical inspection include reviewing the code against a checklist (checking
the consistency of variable names, look for terminating loops, etc.), performing
specific review tasks (such as summarizing the code's purpose, cross-referencing
the code to the technical manual, creating a data dictionary for a given module,
etc.) Rigorous inspection finds 60-65\% of latent defects on average, and often
tops 85\% in defect removal efficiency \citep{Jones2008}. The success rate of
code inspection is generally higher than most forms of testing, which average
between 30 and 35\% for defect removal efficiency \citep{EbertAndJones2009,
Jones2008}. For research software, \citep{KellyAndShepard2000} show a task based
inspection approach can be effective. Task based inspection is an ideal fit with
an issue tracking system, like GitHub.  The review tasks can be issues, so that
they can be easily assigned, monitored and recorded. Potential issues include
assigning junior developers to test getting-started tutorial and installation
instructions.

As indicated in Section~\ref{Sec_CompareMethodologies} some MI projects use
modern code review, via issue tracking and the use of GitHub.  Those MI projects
not incorporating modern code review would likely benefit by adopting it.
Although a rigorous code inspection is likely not worth the required resources,
for critical parts of the code, developers may want to adopt a more rigorous
approach. For instance, developers may drop the modern requirement for
asynchronous review and instead occasionally use synchronous review to help
uncover errors and disseminate best practices throughout the team.  For
instance, teams could periodically meet, either in-person or virtually, and have
junior members walk through their code.  In-person reviews will likely help
realize the benefits of modern code review noticed by
\citet{BirdAndBacchelli2013}: defect detection, knowledge transfer, increased
team awareness, and creation of alternative solutions to problems.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \qref{Q_Maintainability}. Peer review can potentially find misunderstandings
in of how the code implements the required theory, which will improve the
software's correctness (\ppref{P_Correctness}). The benefits of peer review for
addressing pain points can be increased by extending the review from just code,
to also reviewing all software artifacts, including documentation, build
scripts, test cases and the development process itself.

@@

In our ``state of the practice'' assessment exercise for LBM software
\citep{SmithEtAl2022}, we noticed that LBM developers implicitly used
modularization based on the principle of design for change to improve
maintainability (\qref{Q_Maintainability}).  We recommend that MI developers use
the same principle for their modularizations.  Although the advice to modularize
research software to handle complexity is common \citep{WilsonEtAl2014,
StewartEtAl2017, Storer2017}, specific guidelines on how to divide the software
into modules is less prevalent.  Not every decomposition is a good design for
supporting change, as shown by \citet{Parnas1972a}.  A design with low cohesion
and high coupling \citep[p.\ 48]{GhezziEtAl2003} will make change difficult.
Especially in research software, where change is inevitable, designers need to
produce a modularization that supports change. \citep{JungEtAl2022} points out
that ocean modelling software is currently feeling the pain of not emphasizing
modularization in legacy code.

Specific examples of design for change for LBM software \citep{SmithEtAl2022}
include the following:

\begin{itemize}
\item \href{https://github.com/pylbm/pylbm}{pyLBM} has decoupled geometries and
models of their system using abstraction and modularization of the source code,
to make it easy to add new features.  The pyLBM design allows for independent
changes to the geometry and the model.  pyLBM also redeveloped data structures
to ease future change. 
\item \href{https://github.com/CFD-GO/TCLB}{TCLB} \citep{rokicki2016adjoint} is
designed to allow for the addition of some LBM features, but changes to major
aspects of the system would be difficult. For example, ``implementing a new
model will be an easy contribution'', but changes to the ``Cartesian mesh … will
be a nightmare'' \citep{SmithEtAl2022}. The design of TCLB highlights that not
every conceivable change needs to be supported, only the likely changes.  
\end{itemize}

As the LBM examples above illustrate, developers can accomplish design for
change by first identifying likely changes, either implicitly or explicitly, and
second by hiding each likely change behind a well-defined module interface.
This approach mirrors the recommendations from \citet{Parnas1972a}.
\citet{SmithEtAl2022} (under Section ``Documentation as Part of the Development
Process'') lists ideas for how to document the design, including the likely
changes, so that they are more visible to others.

@@

To ensure correctness (\ppref{P_Correctness} and to achieve the quality of
reproducibility (\qref{Q_Reproducibility}), we recommend considering the use of
assurance cases.  \citet{RinehartEtAl2015} defines an assurance case as ``[a]
reasoned and compelling argument, supported by a body of evidence, that a
system, service, or organization will operate as intended for a defined
application in a defined environment.''  An assurance cases provide an organized
and explicit argument that the software and its documentation achieves desired
qualities, such as correctness and reproducibility.  Although assurance cases
have been successfully employed for safety critical systems
\citep{RinehartEtAl2015}, this technique is relatively new for research software
\citep{SmithEtAl2020_AC, Smith2018}.

One way to present an assurance case is through the Goal Structuring Notation
(GSN) \cite{Spriggs2012}, which make arguments clear, easy to read and, hence,
easy to challenge. GSN starts with a Top Goal (Claim), like ``Program X delivers
correct outputs when used for its intended use/purpose in its intended
environment.''  We then decompose this top goal into Sub-Goals, which themselves
may be further decomposed.  The purpose of the decomposition is to take the
abstract higher level goals and bring them down to something concrete that can
be proven.  The decomposition ends with the terminal Sub-Goals that are
supported by Solutions (Evidence). Typical evidence will consist of documents,
expert reviews, test case results, peer review, etc.  Within the GSN framework,
there are also strategy blocks, which describe the rationale for decomposing a
Goal or Sub-Goal into more detailed Sub-Goals. A common tool for creating,
editing, and presenting, a GSN argument is \href{https://astah.net/} {Astah}.

\citet{SmithEtAl2020_AC} shows the example of arguing for the correctness of the
Analysis of Functional NeuroImages (AFNI) package 3dfim+ \citep{Ward2000}.
3dfim+ analyzes the activity of the brain by computing the correlation between
an ideal signal and the measured brain signal for each voxel. The assurance case
for the correctness of 3dfim+ has the top level decomposed into four sub-goals,
as shown in Figure~\ref{TopGoal}.  This example follows the same pattern as used
for medical devices~\cite{Wassyng2015}.  The first sub-goal (GR) argues for the
quality of the documentation of the requirements (SRS).  The second sub-goal
(GD) says that the design complies with the requirements and the third proposes
that the implementation also complies with the requirements.  The fourth
sub-goal (GA) claims that the inputs to 3dfim+ will satisfy the operational
assumptions, since we need valid input to make an argument for the correctness
of the output.

\begin{figure}[!h]
\centering
\includegraphics[width=1.0\textwidth]{./figures/TopGoal.pdf}
\caption{Top Goal of the assurance case and its sub-goals}
\label{TopGoal}
\end{figure}

Preparing an assurance case for the pre-existing 3dfim+ software shows the value
of an assurance cases for research software. Although \citet{SmithEtAl2020}
found no errors in the output of the existing software, the rigour of the
proposed approach did lead to finding ambiguities and omissions in the existing
documentation, such as missing information on the coordinate system convention.
In addition, a potential concern for the software itself was identified from the
GA argument: running the software does not produce any warning about the
obligation of the user to provide data that matches the parametric statistical
model employed for the correlation calculations.

@@

Compare to LBM.

@@

\section*{Acknowledgements}
